A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
minor

@@

\begin{lstlisting}
> eval(2)
[1] 2
> eval(expression(1+2))
[1] 3
> eval(parse(text = "1 + 2"))
[1] 3
\end{lstlisting}

\begin{figure*}[!tb]\centering\includegraphics[width=.8\linewidth]
{pipeline.pdf}\caption{Analysis Pipeline}\label{fig:pipeline}
\end{figure*}

@@

Our infrastructure consists of a dynamic analysis pipeline to record the calls
to \eval and RMarkdown notebooks for data analysis and visualization.

Figure~\ref{fig:pipeline} shows the main steps of the pipeline. We provide the
runtime, input and output data size for each step.

@@

The dynamic analyzer builds upon our dynamic analysis framework, \instrumentr
that we have implemented to enable us to write dynamic analysis logic in R. It
is an R package implemented in 2.5K lines of R and 6K lines of C++ code. It
internally uses a modified R interpreter, \rdyntrace~\cite{oopsla19a}, that
exposes hooks from within the interpreter implementation for events of interest.
\instrumentr serves as an intermediary between \rdyntrace and \evil, it
intercepts the hooks exposed by \rdyntrace and attaches R functions exported by
\evil as callbacks. The \evil callbacks execute on corresponding interpreter
events.

All steps of this pipeline are parallelized using GNU
parallel~\cite{GNUparallel} and orchestrated by GNU make. To schedule and
parallelize extraction and analysis of programs, we use our homegrown library
\runr. Furthermore, \runr gracefully handles and reports failures across
large-scale program runs which greatly aids debugging of the analysis pipeline.
The data extracted by \evil from each program is concatenated, cleaned and
summarized in the post-processing phase by custom R scripts. Finally, the
summarized data is analyzed in RMarkdown notebooks to gather insights. Apart
from the figures, the data points included in the paper are also generated by
RMarkdown notebooks as latex macros.

All runs are done on a single server Intel Xeon 6140, 2.30GHz with 72 cores and
256GB of RAM.

@@

In this study, we report on \eval usage in a corpus of \CorpusCorePackages R core
and \CorpusPackages popular CRAN packages. To quantify the eval use, we run
\CorpusAllProgramsRnd of R programs that are extracted form those packages as
well as from the user-written data-analyses shared on the Kaggle platform. In
this section we describe this corpus and methodology used to assemble it.

@@

We have selected the TOP \CorpusPackages packages based on the number of reverse
dependencies, \ie how many other packages depend on them\footnote{Extracted from
  package metadata using builtin function}. Using this criterion helps us to
choose popular packages which in general should be of a better quality and have
better code coverage and therefore more code to run. The resulting set contains
\CorpusRCodeRnd lines of R code and \CorpusNativeCodeRnd lines of native code
(C/C++/Fortran)\footnote{Excluding comments and blank lines using
  \url{github.com/AlDanial/cloc}}. Figure~\ref{fig:corpus} shows these packages,
the size of the dots reflects the project's size in lines of code including both
R and native code, the x-axis indicates the expression code coverage in percents
and the y-axis gives the number of reverse dependencies in log scale. Dotted
lines indicate means. Packages with over \CorpusEvalsPackageTreshold eval call
sites are annotated.

@@

There are three sources of runnable code in a R package. Next to the package
\emph{tests}, there are also \emph{examples} and \emph{vignettes}. Examples are
R code snippets extracted from package documentation into scripts file.
Vignettes are long-form description of package functionality written using
literate-programming style in \LaTeX\xspace or Rmarkdown with chunks of R code
that can be similarly extracted into R scripts. R provides builtin functions for
tangling the embedded code into files.

>>>>

\begin{figure*}[!t]\centering\includegraphics[width=.8\linewidth]
{pipeline.pdf}\caption{Analysis Pipeline}\label{fig:pipeline}
\end{figure*}

@@

@@

Our infrastructure consists of a dynamic analysis pipeline to record the
calls to \eval and RMarkdown notebooks for data analysis and visualization.
Figure~\ref{fig:pipeline} shows the main steps of the pipeline. We provide
the runtime, input and output data size for each step. All runs are done on
a single server Intel Xeon 6140, 2.30GHz with 72 cores and 256GB of RAM.

@@

The dynamic analyzer builds upon the dynamic analysis framework, \instrumentr
that we have implemented to enable us to write dynamic analysis logic in R. It
is an R package implemented in 2.5K lines of R and 6K lines of C++ code. It
internally uses a modified R interpreter, \rdyntrace~\cite{oopsla19a}, that
exposes hooks from within the interpreter implementation for events of interest.
\instrumentr serves as an intermediary between \rdyntrace and \evil, it
intercepts the hooks exposed by \rdyntrace and attaches R functions exported by
\evil as callbacks. The \evil callbacks execute on corresponding interpreter
events.

All steps of this pipeline are parallelized using GNU
parallel~\cite{GNUparallel} and orchestrated by GNU make. To schedule and
parallelize extraction and analysis of programs, we use the \runr
package. Furthermore, \runr gracefully handles and reports failures across
large-scale program runs which greatly aids debugging of the analysis
pipeline.  The data extracted by \evil from each program is concatenated,
cleaned and summarized in the post-processing phase by custom R
scripts. Finally, the summarized data is analyzed in RMarkdown notebooks to
gather insights. Apart from the figures, the data points included in the
paper are also generated by RMarkdown notebooks as latex macros.

@@

In this study, we report on a corpus of \CorpusCorePackages R core and
\CorpusPackages CRAN packages. We run \CorpusAllProgramsRnd programs
extracted from those packages as well as from the user-written data-analyses
shared on the Kaggle platform. In this section we describe this corpus and
methodology used to assemble it.

@@

We have selected the TOP \CorpusPackages packages based on the number of
reverse dependencies, \ie how many other packages depend on
them\footnote{Extracted from package metadata using a builtin function}.
This criterion helps us find packages which should be better written and
have better code coverage.  The resulting set contains \CorpusRCodeRnd lines
of R code and \CorpusNativeCodeRnd lines of native code
(C/C++/Fortran)\footnote{Excluding comments and blank lines using
  \url{github.com/AlDanial/cloc}}. Figure~\ref{fig:corpus} shows these
packages, the size of the dots reflects the project's size in lines of code.
The x-axis indicates code coverage in percents and the y-axis gives the
number of call sites to \eval in log scale. Dotted lines indicate
means. Packages with over \CorpusEvalsPackageTreshold eval call sites are
named.

@@

There are three sources of runnable code in a R package. Next to the package
\emph{tests}, there are also \emph{examples} and \emph{vignettes}. Examples
are R code snippets extracted from package documentation into scripts files.
Vignettes are long-form description of package functionality written using
literate-programming style in \LaTeX\xspace or Rmarkdown with chunks of R
code that can be similarly extracted into R scripts. R provides builtin
functions for tangling the embedded code into files.
